%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------
\DocumentMetadata{
pdfversion=2.0,  
lang=en-US,   
pdfstandard={ua-2, a-4f},
tagging = on, 
tagging-setup={math/setup={mathml-SE}} ,
}
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage[extreme]{savetrees}
\usepackage[utf8]{inputenc}
\usepackage{newpxtext}
\usepackage{hyperref}
\usepackage{multicol}
\usepackage{physics}
\usepackage{siunitx}

%%%%%%%%% === Document Configuration === %%%%%%%%%%%%%%

\hypersetup{
pdftitle={MATH 340 -- Homework 8 Spring 2026},
pdfauthor={Ignacio Rojas},
pdfkeywords={eighth homework},%
}

\newcommand{\twobyone}[2]{\begin{pmatrix} %% 2 x 1 matrix
   #1 \\ #2 \end{pmatrix}}
\newcommand{\twobytwo}[4]{\begin{pmatrix} %% 2 x 2 matrix
   #1 & #2 \\ #3 & #4 \end{pmatrix}}
\newcommand{\threebyone}[3]{\begin{pmatrix} %% 3 x 1 matrix
   #1 \\ #2 \\ #3 \end{pmatrix}}
\newcommand{\threebythree}[9]{\begin{pmatrix} %% 3 x 3 matrix
   #1 & #2 & #3 \\ #4 & #5 & #6 \\ #7 & #8 & #9 \end{pmatrix}}

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------
\begin{document}

\begin{center}
   {\Large MATH 340 -- Homework 8,\quad Spring 2026}
\end{center}

For this homework, every true/false question is worth one point. You must hand in at least 10 points. When appropriate, you can get another point by giving an example of the fact when it's true or by giving a counterexample when it's false.

\begin{enumerate}

   \item (T) (F) A matrix \(A\) with \(\det(A)\neq 0\) has nonzero eigenvalues.
         %\item (T) (F) If the Wronskian $W[\mathbf{x}_1,\mathbf{x}_2]=0$, then one of $\mathbf{x}_1$, $\mathbf{x}_2$ must not be a solution to $\mathbf{x}'(t)=\mathbf{Ax}(t)$.
   \item (T) (F) If \(\vec{v}_1\) and \(\vec{v}_2\) are eigenvectors of \(A\), then \(\vec{v}_1\) and \(\vec{v}_2\) must be linearly independent.
   \item (T) (F) For an order 2 matrix \(A\) with linearly independent eigenvectors \(\vec{v}_1\) and \(\vec{v}_2\), it is always possible to find an eigenvector \(\vec{v}_3\) which is not a scalar multiple of \(\vec{v}_1\) or a scalar multiple of \(\vec{v}_2\).
         %\item (T) (F) If \(\vec{u}_1\) and \(\vec{u}_2\) are linearly independent constant vectors, then \(\vec{u}_1\exp(tA)\) and \(\vec{u}_2\exp(tA)\) are linearly independent solutions of the ODE system \(\vec{x}'(t)=A\vec{x}(t)\).
   \item (T) (F) If the matrix \(A\) is invertible, then \(\det(A)=0\).
   \item (T) (F) If the order 3 matrix \(A\) has exactly 2 eigenvalues \(\lambda_1\) and \(\lambda_2\), then one of \(\lambda_1\) or \(\lambda_2\) must have an algebraic multiplicity of \(1\).
   \item (T) (F) If the general solution of a 2-dimensional ODE system \(\vec{x}'(t)=A\vec{x}(t)\) is \(c_1\vec{v}+c_2(\vec{v}t+\vec{w})\), then \(A\) has 0 as an eigenvalue with algebraic multiplicity 2.
   \item (T) (F) If \(A\) is an order 2 real matrix with eigenvalues \(1\) and \(-1\), then every solution of the ODE system \(\vec{x}'(t)=A\vec{x}(t)\) that does not start at the origin will diverge.
   \item (T) (F) If \(A\) is an order 3 real matrix, it can only have one complex eigenvalue.
   \item (T) (F) The ODE system \(\begin{cases}
            x'(t)=-x+2y \\
            y'(t)=-6x+6y
         \end{cases}\) has a saddle-point at the origin.
   \item (T) (F) The ODE system \(\begin{cases}
            x'(t)=-2x-y \\
            y'(t)=x-2y
         \end{cases}\) behaves as a center at the origin.
   \item (T) (F) For the ODE system \(\vec{x}'=A\vec{x}\), if \(\Tr(A)=7\) and \(\det(A)=12\), then the origin is a nodal source.
   \item (T) (F) Every order 2, real matrix has \(2\) real eigenvalues.
   \item (T) (F) The zero vector is an eigenvector of every invertible matrix.
   \item (T) (F) A \(2\times 2\) matrix with an eigenvalue of geometric multiplicity \(2\) has the property that the sum of two linearly independent eigenvectors is also an eigenvector.
   \item (T) (F) For a \(2\)-dim'l. ODE system \(\vec{x}'=A\vec{x}\), if \(A\) has real entries and one of its eigenvalues is \(\lambda=-3+4i\) then all of its trajectories spiral towards the origin.
   \item (T) (F) For a \(2\)-dim'l. ODE system \(\vec{x}'=A\vec{x}\), if \(A\) has real entries and one of its eigenvalues is \(\lambda=1-2i\) then all of its trajectories spiral towards the origin.
   \item (T) (F) For a \(2\)-dim'l. ODE system \(\vec{x}'=A\vec{x}\), if \(A\) has real entries and one of its eigenvalues is \(\lambda=-2i\) then all of its trajectories rotate around the origin.
   \item (T) (F) For a \(2\)-dim'l. ODE system \(\vec{x}'=A\vec{x}\), if \(A\) has real entries and \(\det(A)=0\), then the origin is a saddle point.
   \item (T) (F) For a \(2\)-dim'l. ODE system \(\vec{x}'=A\vec{x}\), if \(A\) has real entries and \(\det(A)=0\), then \(\lim_{t\to\infty}\vec{x}(t)=0\) for any solution \(\vec{x}\).
   \item (T) (F) For a \(3\)-dim'l. ODE system \(\vec{x}'=A\vec{x}\), if \(A\) has characteristic polynomial
         \[f_A(\lambda)=(\lambda+2)(\lambda^2+6\lambda+25),\]
         then \(\lim_{t\to\infty}\vec{x}(t)=0\) for any solution \(\vec{x}\).
   \item (T) (F) If \(\lambda\) is a complex eigenvalue of a \(3\times 3\) matrix with real entries, its geometric multiplicity must equal its algebraic multiplicity.
   \item (T) (F) For an ODE system \(\vec{x}'=A\vec{x}\), if \(A\) has real entries and \(\vec{z}(t)\) is a complex valued solution of the system, then \(\Im(\vec{z})\) is a real valued solution of the same system.
   \item (T) (F) If \(2\) is an eigenvalue of an invertible matrix \(A\), then \(\frac12\) is an eigenvalue of \(A^{-1}\).
   \item (T) (F) If \(\lambda\) is an eigenvalue with algebraic multiplicity \(2\), then it surely has two linearly independent eigenvectors.
   \item (T) (F) If a matrix with real entries has a complex eigenvalue \(\lambda\), then \(\bar{\lambda}\) is also an eigenvalue.
   \item (T) (F) If a \(3\times 3\) matrix with real entries has a pair of conjugate eigenvalues, then the third eigenvalue must necessarily be real.

\end{enumerate}

\end{document}