\documentclass[12pt]{memoir}

\def\nsemestre {I}
\def\nterm {Spring}
\def\nyear {2025}
\def\nprofesor {Ignacio Rojas}
\def\nsigla {MATH101}
\def\nsiglahead {Math in the Social Sciences}
\def\nextra {Feedback 1}
\def\nlang {ENG}
\input{../../headerVarillyDiff}
\def\nauthor{}

\begin{document}

Exams are part of college life. Learning from an exam involves reviewing the beliefs, habits, and techniques that contributed to your exam performance. Write a 1-2 paragraph letter to your future self describing what went well on Exam 1 and what could go better. Revisit this letter before Exam 2 to help you prepare. Your letter will only be visible to you and me.\par
Here are some prompts to get you started. You do \textbf{not} need to answer all the questions.

\begin{enumerate}
    \item What study techniques did you use? Of these, what helped you the most?

    \item What resources did you use? Of these, what helped you the most? (Note: resources can include people!)
    
    \item Are there study techniques or resources you wish you had used?
    
    \item On a scale of 1-10, how prepared did you feel before the exam? Did this feel consistent with your exam performance?
    
    \item Think about the questions on the exam. What helped you when you got stuck? What hurt you when you got stuck?
    
    \item Was the format and types of questions on the exam what you expected? If not, what was different?
    
    \item Would you do the same pre-exam or post-exam routine? Why or why not?
    
    \item What advice and encouragement would you give your future self before the second exam?
\end{enumerate}

If you \emph{missed} Exam 1 - write a letter to yourself about what you will do differently to ensure you do not miss Exam 2.\par

I look forward to reading your letters so I can help you prepare for Exam 2.
\end{document} 
