%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------
\DocumentMetadata{
pdfversion=2.0,  
lang=en-US,   
pdfstandard={ua-2, a-4f},
tagging = on, 
tagging-setup={math/setup={mathml-SE}} ,
}
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage[extreme]{savetrees}
\usepackage[utf8]{inputenc}
\usepackage{newpxtext}
\usepackage{hyperref}
\usepackage{multicol}
\usepackage{physics}
\usepackage{siunitx}

%%%%%%%%% === Document Configuration === %%%%%%%%%%%%%%

\hypersetup{
pdftitle={MATH 340 -- Homework 6 Spring 2026},
pdfauthor={Ignacio Rojas},
pdfkeywords={sixth homework},%
}

\newcommand{\twobyone}[2]{\begin{pmatrix} %% 2 x 1 matrix
   #1 \\ #2 \end{pmatrix}}
\newcommand{\twobytwo}[4]{\begin{pmatrix} %% 2 x 2 matrix
   #1 & #2 \\ #3 & #4 \end{pmatrix}}
\newcommand{\threebyone}[3]{\begin{pmatrix} %% 3 x 1 matrix
   #1 \\ #2 \\ #3 \end{pmatrix}}
\newcommand{\threebythree}[9]{\begin{pmatrix} %% 3 x 3 matrix
   #1 & #2 & #3 \\ #4 & #5 & #6 \\ #7 & #8 & #9 \end{pmatrix}}

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------
\begin{document}

\begin{center}
    {\Large MATH 340 -- Homework 6,\quad Spring 2026}
\end{center}

Recall that you must hand in a subset of the problems for which deleting any problem makes the total score less than 10. The maximum possible score on this homework is 10 points. See the syllabus for scoring details. NOTE: All problems from Boyce, Di Prima and Meade are from the \(11^{\text{th}}\)  edition, which is available online: \href{https://www.math.colostate.edu/~liu/MATH340_S26Crd/Textbook_BoyceDiPrimaMeade_11ed.pdf}{https://www.math.colostate.edu/\(\sim\)liu/MATH340\_S26Crd/Textbook\_BoyceDiPrimaMeade\_11ed.pdf}.

\begin{enumerate}
  
  \item (\(\sim\)\textbf{E}) (2) [3 points] Recall the trace-determinant formula for eigenvalues of a \(2\times 2\) matrix \(A= \twobytwo{a}{b}{c}{d}\)
  \[
  \lambda_{1},\lambda_{2}=m\pm\sqrt{m^2-p},\quad\text{where}\quad m=\frac{\Tr(A)}{2}=\frac{a+d}{2},\quad\text{and}\quad p=\det(A)=ad-bc.
  \]
  With this, find the eigenvalues and algebraic multiplicities of the following matrices:
  \[\twobytwo{0}{1}{1}{0},\quad \twobytwo{2}{4}{-1}{-2},\quad \twobytwo{-3}{4}{-1}{1},\quad \twobytwo{-3}{-2}{2}{1},\quad \twobytwo{-3}{2}{-4}{1},\quad \twobytwo{3}{-1}{4}{-2},\quad\twobytwo{5}{-1}{3}{1}\]
  
  \item (\(\sim\)\textbf{E}) (1+) [2 points] Determine the values of the positive constants \(k\) and \(\gamma\) such that the matrices
\[\twobytwo{0}{1}{-k}{1}\quad \text{and} \quad \twobytwo{0}{1}{-1}{-\gamma}\]
have complex eigenvalues with imaginary parts equal to \(3\) and \(1/2\), respectively.

\item (\(\sim\)\textbf{E}) (2) [3 points] Consider the vector function
\[F(x,y)=\twobyone{-(x+y)(1-x-y)}{x(2+y)}.\]
Find the Jacobian of \(F\) and evaluate it at \((0,0)\) and \((0,1)\). Determine the eigenvalues for both cases.

\item (1+) [2 points] In order to send an undetected message to an agent in the field, each letter in the message is replaced by the
number of its position in the alphabet and that number is entered in a matrix \(M\). Thus, for example, "READ" becomes the matrix \(M=\twobytwo{18}{5}{1}{4}\).\par
In order to further avoid detection, each message with four letters is sent to the agent encoded as \(MC\), where \(C=\twobytwo{2}{-1}{1}{1}\). If the agent receives the matrix \(\twobytwo{51}{-3}{31}{-8}\), then what was the original message?

\item (1+) [2 points] Let \(0<t<\pi\). The matrix \(\twobytwo{\cos(t)}{-\sin(t)}{\sin(t)}{\cos(t)}\) has eigenvalues \(\lambda_1,\lambda_2\). For what value of \(t\), does it hold that \(\lambda_1+\lambda_2=1\)?

\item (1+) [2 points] Boyce, di Prima, and Meade chapter 7 section 2 problem 17.
  
\item (2) [2 points] Boyce, di Prima, and Meade chapter 7 section 2 problem 18.

\item (3-) [8 points] Find the eigenvalues and eigenvectors of the following \(3\times 3\) matrices
\[
\threebythree{0}{0}{-2}{0}{2}{2}{-2}{2}{1},\quad\threebythree{-2}{0}{2}{0}{0}{2}{2}{2}{1},\quad\threebythree{-1}{-4}{-4}{-4}{-3}{0}{-4}{0}{1}.
\]

%{{1,0,-1},{0,1,-1},{0,0,-1}}*{{1,0,0},{0,2,1},{0,0,2}}*{{1,0,-1},{0,1,-1},{0,0,-1}}^(-1)
%{{0,-1,1},{-1,1,0},{1,0,0}}*{{2,0,0},{0,1,1},{0,0,1}}*{{0,-1,1},{-1,1,0},{1,0,0}}^(-1)
%{{1,-1,1},{-1,1,0},{1,0,0}}*{{1,0,0},{0,2,1},{0,0,2}}*{{1,-1,1},{-1,1,0},{1,0,0}}^(-1)
\item (3-) [8 points] Write down the Jordan decomposition of the following matrices
\[
\threebythree{1}{0}{1}{0}{2}{-1}{0}{0}{2},\quad\threebythree{0}{-1}{-1}{1}{2}{0}{0}{0}{2},\quad\threebythree{1}{-1}{-1}{1}{3}{1}{0}{0}{1}
\]
\item (1-) [1 point] If \(f\) is a linear vector function such that 
\[f(1,1)=1,\quad\text{and}\quad f(-1,0)=2,\]
how much is \(f(3,5)\)?

\item (1-) [1 point] If \(A\) is a \(3\times 3\) matrix such that 
\[
A\threebyone{0}{1}{2}=\threebyone{1}{0}{0},\quad\text{and}\quad A\threebyone{3}{4}{5}=\threebyone{0}{1}{0},
\]
what is the value of \(A\threebyone{6}{7}{8}\)?

%1
\item (2) [3 points] For which value of \(b\) does the matrix
\[A=\threebythree{0}{3}{1}{0}{0}{-12}{0}{3b}{12}\]
have an eigenvalue of algebraic multiplicity \(2\)?

\item (2+) [4 points] Supppose \(a>0\). If 
\[A=\threebythree{a}{0}{0}{-a}{a}{-a}{a}{0}{2a},\]
find the eigenvector(s) corresponding to \(A\)'s smallest eigenvalue.

\item (1+) [2 points] Let \(a,b>0\) and consider the matrix
\[
\threebythree{a}{b-a}{a-b}{0}{a}{0}{0}{a-b}{b}
\]
with characteristic polynomial \(p(\lambda)=(\lambda-a)^2(\lambda-b)\). The eigenspace associated to \(\lambda=a\) is spanned by the vectors 
\[\vec{v}_1=\threebyone{a}{b}{b},\quad\text{and}\quad\vec{v}_2=\threebyone{a}{-b}{-b}.\]
With this information, is it possible to conclude that the matrix is diagonalizable? [Hint: Recall the AMGM inequality.]
\end{enumerate}

\end{document}